Thesis Draft

### 2017-09-28 thesis/draft.Rmd
corrected NA fix code, where true NA was incorrectly made 0 values

dfa$DayType <-       ifelse(dfa$Day == "Sunday", "Sunday",
                     ifelse(dfa$Day == "Saturday", "Saturday",
                    ifelse(dfa$Day == "Monday", "Monday",
                     ifelse(dfa$Day == "Friday", "Friday",
                                  "Midweek"))))
dfa$DayType[dfa$IsHDay == 1] <- "Holiday"
dfa$HDay <- ifelse(dfa$IsHDay == 1, "Holiday", dfa$Day)
original_df <- dfa
dfa <- filter(dfa, Date_Time < "2017-01-01")
```
## Background

@@

```{r}
cl <- makeCluster(2)
registerDoSNOW(cl)
models <- foreach(i = unique(ped_data$Sensor_ID)) %dopar% {
  tr_dat <- data.frame(dfa [, i],  HDay = dfa$HDay, Time = dfa$Time)
  colnames(tr_dat)[1] <- "Hourly_Counts"
  model <- glm(Hourly_Counts ~ HDay*Time, data = tr_dat, family = quasipoisson())
  model <- stripGlmLR(model)
  return(model)
}

# ped_models[[31]]$xlevels <- ped_models[[1]]$xlevels

@@

run_lengths <- foreach(i = unique(ped_data$Sensor)) %do%
  {
    dfa[is.na(dfa[, i]), i] <- 0
    na_length_check <- rle(as.numeric(unlist(dfa[,i])))
    maybeNA <- rep((na_length_check$lengths > threshold_miss_hrs), times = na_length_check$lengths)
    rl <- dfa[, i] %>% unlist %>% rle()
    dfa[maybeNA, i] <- NA
    return(rl)

    }
run_lengths %>% lapply(., `[[`, 1) %>%
    unlist -> sequences

@@

```{r alg_glm}
models <- list()
fitted_data <- list()
used_sensors <- list()
dfa2 <- dfa
p <- progress_estimated(length(unique(ped_data$Sensor)))
for(i in small_miss_sensors){
  p$tick()$print()
  options(na.action = na.omit)
  # if(max(na_length_check$lengths) > max_miss_hrs) 
  # {
  #   model <- NULL
  # }

@@

For the purposes of evaluating the imputation method at a location with a small proportion of missing values, we will first look at the counts from Southern Cross Station. Firstly, we evaluate the goodness of fit of the predictions made within the training period (01/01/2014 to 31/12/2016).

```{r}
ggplot(data = NULL, aes(x = original_df %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Southern Cross Station"],
           y = fitted_data$`Southern Cross Station`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = brewer.pal(3, 'Dark2')[1],
                       alpha = 0.7, size = 1.5,) +
           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
                       colour = brewer.pal(3, 'Dark2')[2]) +
           theme_minimal() + theme(aspect.ratio = 1) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Actual against Predicted Counts at Southern Cross Station")

@@

>>>>

dfa$DayType <-       ifelse(dfa$Day == "Sunday", "Sunday",
                     ifelse(dfa$Day == "Saturday", "Saturday",
                    ifelse(dfa$Day == "Monday", "Monday",
                     ifelse(dfa$Day == "Friday", "Friday",
                                  "Midweek"))))
dfa$DayType[dfa$IsHDay == 1] <- "Holiday"
dfa$HDay <- ifelse(dfa$IsHDay == 1, "Holiday", dfa$Day)
original_df <- dfa
dfa <- filter(dfa, Date_Time < "2017-01-01")
ori_2 <- filter(original_df, Date_Time < "2017-01-01")
```
## Background

@@

```{r}
cl <- makeCluster(2)
registerDoSNOW(cl)
fitted_base <- list()
models_base <- foreach(i = unique(ped_data$Sensor_ID)) %dopar% {
    tr_dat <- data.frame(dfa[, i],  HDay = dfa$HDay, Time = dfa$Time)
    colnames(tr_dat)[1] <- "Hourly_Counts"
    model <- glm(Hourly_Counts ~ HDay*Time, data = tr_dat, family = quasipoisson())
    return(stripGlmLR(model))
  }
names(models_base) <- unique(ped_data$Sensor_ID)
# ped_models[[31]]$xlevels <- ped_models[[1]]$xlevels
for (i in unique(ped_data$Sensor)){  
  tr_dat <- data.frame(dfa [, i],  HDay = dfa$HDay, Time = dfa$Time)
  options(na.action = na.pass)
  fitted_base[[i]] <- predict.glm(models_base[[i]], newdata = tr_dat, type = "response")
  options(na.action = na.omit)
  }

@@

run_lengths <- list()

for(i in unique(ped_data$Sensor)){
    dfa[is.na(dfa[, i]), i] <- 0
    na_length_check <- rle(as.numeric(unlist(dfa[,i])))
    maybeNA <- rep((na_length_check$lengths > threshold_miss_hrs), times = na_length_check$lengths)
    rl <- dfa[, i] %>% unlist %>% rle()
    dfa[maybeNA, i] <- NA
    defsNA <- is.na(ori_2[, i])
    dfa[defsNA, i] <- NA
    run_lengths[[i]] <- rl
  }

run_lengths %>% lapply(., `[[`, 1) %>%
    unlist -> sequences

@@

```{r alg_glm}
models <- list()
fitted_data <- list()
dfa2 <- dfa
p <- progress_estimated(length(unique(ped_data$Sensor)))
for(i in small_miss_sensors){
  p$tick()$print()
  options(na.action = na.omit)
  # if(max(na_length_check$lengths) > max_miss_hrs) 
  # {
  #   model <- NULL
  # }

@@

Firstly, we evaluate the initial model used to impute:
$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$

Note, this is also before the adjustments made to treat suspiciously long sequences of 0 values as `NA`.

For the purposes of evaluating the imputation method at a location with a small proportion of missing values, we will first look at the counts from Southern Cross Station. At this sensor, the proportion of missing values (adjusted) is `r round(na_prop["Southern Cross Station"]*100,2)`Firstly, we evaluate the goodness of fit of the predictions made within the training period (01/01/2014 to 31/12/2016).

```{r}
ggplot(data = NULL, aes(x = original_df %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Southern Cross Station"],
           y = fitted_base$`Southern Cross Station`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = brewer.pal(3, 'Dark2')[1],
                       alpha = 0.7, size = 1.5,) +
           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
                       colour = brewer.pal(3, 'Dark2')[2]) +
           theme_minimal() + theme(aspect.ratio = 1) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Actual against Predicted Counts at Southern Cross Station")

@@

However, if we try to use this same model specification at sensor locations with large proportions of missing values, the fit is poor. We use the sensor at Australia on Collins to 
```{r}
options(na.action = na.omit)
ggplot(data = NULL, aes(x = original_df[-(is.na(original_df$`Australia on Collins`)), ] %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Australia on Collins"],
           y = fitted_base$`Australia on Collins`[-(is.na(original_df$`Australia on Collins`))])) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = brewer.pal(3, 'Dark2')[1],
                       alpha = 0.7, size = 1.5,) +
           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
                       colour = brewer.pal(3, 'Dark2')[2]) +
           theme_minimal() + theme(aspect.ratio = 1) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Actual against Predicted Counts at Australia on Collins")

```
We can see that the model is heavily biased by the false 0 values, causing the predicted values to be underestimate (slope of the fitted line is $<1$).
